---
title: "R Notebook"
output: html_notebook
---

Load all the libraries first:
```{r}
library(dplyr)
library(adaptMCMC)
library(ggplot2)
library(reshape2)
library(GGally)
library(scales)
```

Load the ray-tracer function from the shared object Fortran:

```{r}
setwd('/home/geoanton/RayTracerFortran/RayTracer/')
dyn.load("./subroutineR.so")




#srcd <- src.df$z
#srco <- sqrt(src.df$x**2+src.df$y**2)



```

Set up the subsurface layered model:

```{r}
# The velocity model : v -vels, d - depths of the interfaces
v = c(2100,3270,4070)
d= c(2000, 4000)
NLayers <- length(d)

```

Set up the source-receiver pairs :
```{r}
#src.df <- read.csv('/home/geoanton/WI_modeling/eqdf.csv',header=T)
set.seed(11)
nstation <- 50
src.df <- data.frame(x = runif(n=nstation,min=300,max = 8000),y = runif(n=nstation,min=300,max = 8000),z = runif(n=nstation,min=1500,max = 6000) )
srcd <- src.df$z
srco <- round(sqrt(src.df$x**2+src.df$y**2))
write.csv(src.df,file = "eqdf2.csv")
NSrc <- length(srcd)


```

Run the forward problem :
```{r}
vv <- c(2100.00,3270.00)
dd <- c(2000.00,416.57)
R <- 2940
xx <- seq(1e-4,1/max(vv)-1e-10 , length.out = 500)
cf <- sapply(xx, function(x){ R - sum(dd*vv*x/sqrt(1-(x**2)*vv**2))})

qplot(x=xx,y=cf,geom="line")

```

```{r, include=TRUE}
timeP <- rep(0,NSrc)
#capture.output(
t1 <- Sys.time()
for (i in seq(150))
{
timeP.list <-  .Fortran(
                  "dff" ,vels=as.numeric(v),depths=as.numeric(d),NLayers=as.integer(NLayers),
                  src_offset=as.numeric(srco),src_depth=as.numeric(srcd),NSrc=as.integer(NSrc),
                  timeP = as.numeric(timeP)
                  )
timeP <- timeP.list$timeP
}
t2 <- Sys.time()
#file = "FortranOutput.txt"
#)

#data = timeP$timeP
```


```{r}
data[is.na(data)] = 200
df <- data.frame(x = srco,y=srcd,c = data)

ggplot(data =df,aes(x=x,y=y,colour = c)) + geom_point(size = 5)
```



```{r}
getData <- function(t=seq(1,500),kx = 0.008,k=100,sY=5,sX=10,sigma = 7)
{
  noise =rnorm(length(t) , mean=0,sd=sigma)
  y = k*exp(-kx*(t-sX)) + sY + noise
  return(y)
}

priorTest <- function(x)
# x = c(k,kx,sX,sY)
{
p.k = dunif(x[1], min = 0,max=200,log = T)
p.kx = dunif(x[2], min =0, max = 0.5,log = T)
p.sX = dunif(x[3], min =-5, max=20,log = T)
p.sY = dunif(x[4], min =-5, max=10,log = T)
p.all = p.k+p.kx+p.sY+p.sX
return(p.all)
}
likelihood <- function(true.data,pred.data,sigma.inv.data,sigma.det)
{
  dr <-  pred.data - true.data
  #print(dr)

  N <- length(dr)
  res.norm <- dr %*% sigma.inv.data %*% dr
  print(res.norm)
  ll <- -N*log((sqrt(2*pi))) - sigma.det - 0.5*res.norm
  return(ll)
}
posterior <- function(x,t,true.data,sigma.inv.data,sigma.det)
{
  pred.data <- getData(t=t,k=x[1],kx=x[2],sX=x[3],sY=x[4],sigma=0)
  ll <- likelihood(true.data = true.data,pred.data = pred.data,
                   sigma.inv.data=sigma.inv.data,sigma.det = sigma.det)
  lpr <- priorTest(x)
  
  lpost <- ll+lpr
  
  
}

t = seq(1,500,by=5)
s=20
y <- getData(t,sigma = s)
true.data <- y
qplot(t,y)
sigma.data <- diag(rep((1.2*s)**2,length(t)))
sigma.inv.data <- solve(sigma.data)
sigma.det <- determinant(sigma.data)$modulus
#print(sigma.data)
init.val <-  c(50,0.005,12,0)
scale.x <- c(15,0.002,2,1)**2
pred.data <- getData(t=t,k=init.val[1],kx=init.val[2],sX=init.val[3],sY=init.val[4],sigma=0)
mcchain <- MCMC(p = posterior, n = 20000, init = init.val, scale = scale.x,
                adapt = TRUE, acc.rate = 0.3, gamma = 0.6,
                list = TRUE, showProgressBar=TRUE, n.start = 300, t,true.data,sigma.inv.data,sigma.det)
#qplot(t,pred.data)
#print(pred.data)


```

Plots below:
```{r}
samples <-data.frame(k = mcchain$samples[,1],kx= mcchain$samples[,2],sX = mcchain$samples[,3], sY = mcchain$samples[,4],id = seq(nrow(mcchain$samples)) )

p <- samples %>% select(-id) %>% ggpairs()

samples.melted <- (melt(samples,measure.vars = c('k','kx','sX','sY') ,id.vars = "id" ))
print(mcchain$acceptance.rate)
print(p)
pc <- ggplot(data=samples.melted,aes(x=id,y=value)) + facet_grid(variable~.,scales = "free_y") + geom_line()
print(pc)

```

To coda object
```{r}

coda.sample <- convert.to.coda(mcchain)
plot(coda.sample)

```

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).
