---
title: "RayTracerR Data Saving Test"
output: html_notebook
---

# Foreword 

This notebook is used for developing the function for saving the data in the appropriate format for Jan's Parallel Tempered MCMC. Therefore, it mostly goes over the same procedure that is defined in `raytraceR.R`, but saves data after running the simulation.

# Run the simulation
## Prepare the libraries

```{r message=FALSE, warning=FALSE}
setwd('/home/geoanton/RayTracerFortran/RayTracer/')
library(dplyr)
library(hexbin)
library(ggplot2)
library(reshape2)
library(pbapply)
library(gridExtra)
# define jet colormap
jet.colors <- colorRampPalette(c("#00007F", "blue", "#007FFF", "cyan", "#7FFF7F", "yellow", "#FF7F00", "red", "#7F0000"))

dyn.load("./subroutineR-quiet.so")
base.name <- 'test_1'
source('~/RayTracerFortran/RayTracer/MiscFunctions.R')
```
## Setup source offsets and depths

From file:
```{r}
src.df <- read.csv('/home/geoanton/RayTracerFortran/RayTracer/eqdf.csv',header=T)
srcd <- src.df$z
srco <- sqrt(src.df$x**2+src.df$y**2)
NSrc <- length(srcd)
src.to.write <- tibble(offset = srco,depth=srcd)
print(src.to.write)

```

or randomly..:

```{r}
NSrc <- 5*20*2
set.seed(12)
src.df <- tibble(z=runif(NSrc,min=3050,max=4200),x=runif(NSrc,min=10,max=6500),y=runif(NSrc,min=10,max=4500))
srcd <- src.df$z
srco <- sqrt(src.df$x**2+src.df$y**2)
src.to.write <- tibble(offset = srco,depth=srcd)
```

## Setup velocity model

Simple three-layered model here:

```{r fig.height=5, fig.width=2}
#v <- c(2501,3800,4900,4000,4000)  
#d <- c(2000,3000,3500, 4000)   # That's not thicknesses, that's depth to interfaces !

v <- c(3100,2400,3730,4200,5800,4420,6200)
d <- c(1000,1500,2000,2500,3260,3960)   # That's not thicknesses, that's depth to interfaces !

SdRT = 0.016
NLayers <- length(d)
df.vd <- drawProfiles(c(v,0,d),NLayers)
df.vd <- df.vd %>% arrange(d)
true.plot <- ggplot(data = df.vd,aes(x=v,y=d))  + geom_path(col='red')+ labs(x = "Vel [m/s]",y = "Depth [m]")+ 
                        scale_y_reverse(limits = c(7990,10))
              
print(true.plot)
NLayersMax <- 10
```



## Run forward model and save the 

```{r}
kd <- 10 # keep rays in rays.dat ???
timeP <- rep(0,NSrc)
timeP <- .Fortran("dff" ,vels=as.numeric(v),depths=as.numeric(d),NLayers=as.integer(NLayers),
                  src_offset=as.numeric(srco),src_depth=as.numeric(srcd),NSrc=as.integer(NSrc),
                  timeP = as.numeric(timeP),keep_delta=as.integer(kd))
print(timeP$timeP)
```
and save its results:
```{r}
# Save the travel times into a one column file:
timeP$timeP <- timeP$timeP + rnorm(NSrc,mean=0,sd=SdRT)
write.table(tibble(t=timeP$timeP),file = sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_RT.txt',base.name),sep = ' ',row.names = F,col.names = F)
# Save the source offsets and depths
write.table(src.to.write,file = sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_src_data.txt',base.name),sep = ' ',row.names = F,col.names = F)
```


## Setup the parameter.dat file for MCMC

Here we set the parameters of the MCMC simulations instead of manually correcting them :

```{r}
# NFPMX+1+1
# NFPMX = NPL * NLMX
param.list <- list(NPL = 2, ISMPPRIOR = 0, IEXCHANGE = 1, NDAT_RT = NSrc,
            #       NMODE = 1, NSRC=NSrc,NLMX = NLayers+1, NLMN = NLayers+1, ICHAINTHIN = 5,
                   NMODE = 1, NSRC=NSrc,NLMX = 10, NLMN = 1, ICHAINTHIN = 5,
                   NKEEP=100, NPTCHAINS = 1, dTlog=1.2, lambda = 3.01, hmx = 10000.1, hmn = 100.1,dVs = 0.1, dVpVs=0.1,
                   VpVsmin = 1.5, VpVsmax = 1.8,sdmin = c(0.001,0.001),sdmax = c(0.07,0.07),
                   # MAP below:
                   MAP.k=NLayers+1, MAP.vel = v, MAP.d = c(0,d), MAP.sd = SdRT)
makeParameterFile(param.list = param.list,output.file = '~/RayTracerFortran/RayTracer/test_1/test_1_parameter.dat') ## Create  a parameter file
makeMAPFile(param.list = param.list,output.file = '~/RayTracerFortran/RayTracer/test_1/test_1_map.dat') ## Create  a parameter file
```


## Check the ray results if debugging was switched on
```{r fig.height=10, fig.width=10}
rays.df <- read.rays(nrays=NSrc,timeP = timeP$timeP)
p <- ggplot(data=rays.df,aes(x=delta.cum,y=depth.cum,label = ray.label,col = factor(ray))) + geom_line() + scale_y_reverse()+geom_point()+geom_hline(yintercept = c(0,d),linetype=2)+geom_label()
ggsave(filename = './pngs/ray_example.png',p,width = 7,height = 4)
ggsave(sprintf('/home/geoanton/RayTracerFortran/RayTracer/%s/%s_rays.png',base.name,base.name ),p,width = 12,height=8)

print(p)
print(rays.df)
```

## Plot the sampling from the voro_sample file
We need to follow the notation given in prmjh_temper.f90 and read in the numbers accordingly :
`sample(ikeep,:) =  (/ objm(ic)%logL, objm(ic)%logPr, objm(ic)%tcmp, REAL(objm(ic)%k,RP), & ! 4 parameters
                           tmpvoro,objm(ic)%sdparRT,objm(ic)%arpar,objm(ic)%arparRT, &
                           REAL(iaccept,RP)/REAL(iaccept+ireject,RP),REAL(objm(ic)%iaccept_bd,RP),&
                           REAL(objm(ic)%ireject_bd,RP),REAL(objm(ic)%iaccept_bds,RP),REAL(ic,RP),REAL(rank,RP) /)`
Here we setup the names of the columns, as well as read the file with the sample table:

```{r}
col.names <- c('LogL','LogPrior','tcmp','k')
k <- NLayers+1 # by convention of Jan's RJMCMC
NParamPerLayer <- 2
NVoro <- param.list$NLMX*NParamPerLayer
col.names.voro <- rep(NA,NVoro)
# Make Voro Column names - this will be my velocity profiles
vel.names <- paste0(('V'),seq(param.list$NLMX))
depth.names <- paste0(('Z'),seq(param.list$NLMX))
col.names.voro[seq(1,NVoro-1,by=2)] <- depth.names
col.names.voro[seq(2,NVoro,by=2)] <- vel.names
after.voro.names <- c('SdRT','ArPar','ArParRT','acc_ratio','iacc_bd','irej_bd','iacc_bds','ic','rank')
all.names <- c(col.names,col.names.voro,after.voro.names)
var.to.keep <- c(col.names.voro,'SdRT')
# Prepare the true.model:
true.model = tibble(variable =c(vel.names[1:param.list$MAP.k],depth.names[1:param.list$MAP.k],"SdRT"),value = c(v,c(0,d),SdRT))
# Read the chain from the basename_voro_sample.txt
filename <- '~/RayTracerFortran/RayTracer/test_1/test_1_voro_sample.txt'
chain.df <- read.table(filename) %>% tbl_df() %>% mutate(iter = row_number())
colnames(chain.df)[1:length(all.names)] <- all.names 
```

# Marginal distributions
In order to look at the distributions for each parameter, we need to melt the table:
```{r fig.height=4, fig.width=12}

#chain.df <- chain.df %>% filter(Z2>0)
chain.melted <- melt(chain.df,id.vars = 'iter') %>% filter(variable %in% var.to.keep) %>% tbl_df() %>% mutate(value = as.numeric(value))
p <- ggplot(data = chain.melted,aes(x=value,group = variable)) + geom_histogram(bins=100) + facet_wrap(~variable,scales = 'free',ncol = 4) + geom_vline(aes(xintercept=value),data = true.model,col = 'salmon',size=1.2)
p.traces <- ggplot(data = chain.melted,aes(x=iter,group = variable,y=value)) + geom_line() + facet_wrap(~variable,scales = 'free',ncol = 1)
print(p)

print(p.traces)
ggsave(sprintf('/home/geoanton/RayTracerFortran/RayTracer/%s/%s_marginals.png',base.name,base.name),p,width = 12,height=4)
ggsave(sprintf('/home/geoanton/RayTracerFortran/RayTracer/%s/%s_traces.png',base.name,base.name ),p.traces,width = 12,height=k*2)


```

# Velocity profiles from the chains (FIXED-D):

```{r}
# Profile drawing function:
thin=10
burnin =1000
samples <- chain.df %>% select(c(vel.names,depth.names))
samples.thinned <- as.matrix(samples[seq(burnin,nrow(samples),by=thin),1:ncol(samples)])
profs <-do.call("rbind",pblapply(seq(nrow(samples.thinned)), function(i) {df <- drawProfiles(samples.thinned[i,],NLayers);df$id=i;return(tbl_df(df))},cl=4)) %>% tbl_df()
prof.plot <- ggplot(data = profs,aes(x=v,y=d)) + geom_bin2d(bins =100) + labs(x = "Vel [m/s]",y = "Depth [m]")+ 
              theme(panel.background = element_rect(fill=alpha("#00007F",0.9), colour=alpha("#00007F",0.9)),panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
              scale_y_reverse(limits = c(7990,10))+
              scale_fill_gradientn(colors = jet.colors(500))
newdat <- ggplot_build(prof.plot)$data[[1]]
# dark blue = #132B43
profs.p <- profs  %>% arrange(id,d)
prof.plot.scatter <- ggplot(data = profs.p,aes(x=v,y=d,colour=id)) + geom_path(size=2,colour="blue",alpha=0.02)+ scale_y_reverse(limits = c(7990,10)) + labs(x = "Vel [m/s]",y = "Depth [m]") 
ggsave( sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_scatter.png',base.name),prof.plot.scatter,width = 5,height=12)
# Normalize the profile of the velocities within ybins !
# otherwise prof.plot can be saved 
newdat.norm <- newdat %>% group_by(ybin) %>% mutate(val.norm = value / max(value))
p.norm<- ggplot(data = newdat.norm,aes(x=x,y=y,fill=val.norm)) + geom_raster() + labs(x = "Vel [m/s]",y = "Depth [m]")+ 
              theme(panel.background = element_rect(fill=alpha("#00007F",0.5), colour=alpha("#00007F",0.02)),panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
              scale_fill_gradientn(colors = jet.colors(500))+ geom_path(aes(x=v,y=-d), data = df.vd,inherit.aes = F,col='black',linetype = 2,size=2) +
              geom_point(aes(x=max(profs.p$v),y=-srcd), data = src.to.write,inherit.aes = F,col='white',pch=2,size=2)
ggsave( sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_profiles.png',base.name),p.norm,width = 5,height=12)
```

# Velocity profiles from the chains (TRANS-D):


```{r}
# Profile drawing function:
thin=10
burnin =1000
samples <- chain.df %>% select(c(vel.names,depth.names))
k.layer <- chain.df %>% select(k)
samples.thinned <- as.matrix(samples[seq(burnin,nrow(samples),by=thin),1:ncol(samples)])
k.layer <- as.matrix(k.layer[seq(burnin,nrow(samples),by=thin),])



profs <-do.call("rbind",pblapply(seq(nrow(samples.thinned)), function(i) {df <- drawProfiles.TD(samples.thinned[i,],k.layer[i],param.list$NLMX);df$id=i;return(tbl_df(df))},cl=4)) %>% tbl_df()

prof.plot <- ggplot(data = profs,aes(x=v,y=d)) + geom_bin2d(bins =100) + labs(x = "Vel [m/s]",y = "Depth [m]")+ 
              theme(panel.background = element_rect(fill=alpha("#00007F",0.9), colour=alpha("#00007F",0.9)),panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
              scale_y_reverse(limits = c(7990,10))+
              scale_fill_gradientn(colors = jet.colors(500))

newdat <- ggplot_build(prof.plot)$data[[1]]
# dark blue = #132B43
profs.p <- profs  %>% arrange(id,d)
prof.plot.scatter <- ggplot(data = profs.p,aes(x=v,y=d,colour=id)) + geom_path(size=2,colour="blue",alpha=0.02)+ scale_y_reverse(limits = c(7990,10)) + labs(x = "Vel [m/s]",y = "Depth [m]") 
ggsave( sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_scatter-transD.png',base.name),prof.plot.scatter,width = 5,height=12)
# Normalize the profile of the velocities within ybins !
# otherwise prof.plot can be saved 
newdat.norm <- newdat %>% group_by(ybin) %>% mutate(val.norm = value / max(value))
p.norm<- ggplot(data = newdat.norm,aes(x=x,y=y,fill=val.norm)) + geom_raster() + labs(x = "Vel [m/s]",y = "Depth [m]")+ 
              theme(panel.background = element_rect(fill=alpha("#00007F",0.5), colour=alpha("#00007F",0.02)),panel.grid.major = element_blank(), panel.grid.minor = element_blank())+
              scale_fill_gradientn(colors = jet.colors(500))+ geom_path(aes(x=v,y=-d), data = df.vd,inherit.aes = F,col='black',linetype = 2,size=2) +
              geom_point(aes(x=max(profs.p$v),y=-srcd), data = src.to.write,inherit.aes = F,col='white',pch=2,size=2)
p.k <- ggplot(data = chain.df,aes(x=factor(k) )) + geom_histogram(stat='count') + geom_vline(xintercept = param.list$MAP.k,col='salmon')

p.grid <- arrangeGrob(p.norm,p.k,heights = c(0.8,0.2),nrow = 2)
ggsave( sprintf('/home/geoanton/RayTracerFortran/RayTracer/test_1/%s_profiles-transD.png',base.name),p.grid,width = 5,height=12)

```





Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Ctrl+Alt+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

The preview shows you a rendered HTML copy of the contents of the editor. Consequently, unlike *Knit*, *Preview* does not run any R code chunks. Instead, the output of the chunk when it was last run in the editor is displayed.
